\pkg{SVMBridge}: Using SVM Solvers from R

### 2015-09-15 paper/article.tex
update paper

\documentclass[article]{jss}

@@

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Ayd\i n Demircio\u{g}lu\\
  Institut f\"ur Neuroinformatik\\
  Ruhr-Universit\"at Bochum\\
  44790 Bochum
  E-mail: \email{aydin.demircioglu@ini.rub.de}\\
  URL: \url{http://www.ini.rub.de/PEOPLE/aydind/}
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/512/507-7103
%% Fax: +43/512/507-2851

@@

Support Vector machines in general.

@@

Other SVM solver, like the BudgetedSVM package, which contains
BSGD and LLSVM, or SVMperf cannot be called directly from R.
Instead, a system command has to be issued.
To ease this calls, the \pkg{SVMBridge} package has been developed.

@@

\section{Examples}

@@

@@

Most SVM solver work with the sparse data format.
This format consists of:
Each line (delimited by a CR/LF) is given by a label
and the non-zero components of the data point,
e.g. to encode the vector six dimensional vector 
$0 0 0 0.2 0 1.4$ belonging to class $3$, the sparse 
data format would contain the line {3 4:0.2 6:1.4}.

Although reading these files into R is possible via either
the kernlab or the e1071 routines, both only provide R solutions,
and therefore suffer from suboptimal performance.
The SVMBridge package provides a simple sparse data format 
reading and writing, which are implemented in C++ and
therefore are nearly two orders of magnitudes faster than
the corresponding e1071 routines.
Notice, that currently only dense matrices are supported,
TODO: work around this with the Sparse Matrix Package.

@@

@@

@@

@@

>>>>

\documentclass[article, shortnames]{jss}

% TODO:
% C++/R -> prolang{.}
% SVMBridge -> pkg{.}

\usepackage{amsmath, amssymb}
\usepackage{mathtools}

\newcommand{\sign}{\mathop{\mathrm{sign}}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Order}{\mathcal{O}}

@@

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Ayd\i n Demircio\u{g}lu, Hanna Houphouet, Tobias Glasmachers\\
  Institut f\"ur Neuroinformatik\\
  Ruhr-Universit\"at Bochum\\
  44790 Bochum\\
  Germany\\
  E-mail: \email{\{aydin.demircioglu, hanna.houphouet, tobias.glasmachers\}@ini.rub.de}\\
  URL: \url{http://www.ini.rub.de}\\
  \\

  Daniel Horn, Claus Weihs\\
  Fakult\"at Statistik\\
  Technische Universit\"at Dortmund\\
  44221 Dortmund\\
  Germany \\ 
  E-mail: \email{\{dhorn, bischl, weihs\}@statistik.tu-dortmund.de} \\
  URL: \url{https://www.statistik.tu-dortmund.de/computationalstats.html}\\
\\

  Bernd Bischl\\
Institut f\"ur Statistik \\
Ludwig-Maximilians-Universit\"at M\"unchen\\
80539 M\"unchen\\
  Germany \\
E-mail: \email{bernd.bischl@stat.uni-muenchen.de}\\
URL: \url{http://www.statistik.lmu.de/~bischl}
}

%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/512/507-7103
%% Fax: +43/512/507-2851

@@

Support Vector Machines (SVM) are linear classifiers
that try to maxime the margin of a binary problem \citep{cortes1995support}.
As SVMs are only linear classifiers, a kernelized version
is used for more complex data. Basically, they work by
solving the  following convex problem:
\begin{equation}
        \min_{w \in \mathcal{H}, b \in \R} \quad \frac{1}{2} ||w||^2 + C \cdot \sum_{i=1}^n \max \Big( 0, 1 - y_i \big( \langle w, \varphi(x_i) \rangle_{\mathcal{H}} + b \big) \Big).
        \label{SVMproblemPrimal}
\end{equation}
Here 
$\mathcal{D}=\{(x_1, y_1), \dots, (x_n, y_n)\} \in (X \times \{ \pm 1\})^n$
is the labeled data,
and  $\varphi : X \to \mathcal{H}$ is a feature map into a reproducing
kernel Hilbert space $\mathcal{H}$, corresponding to a positive definite
(Mercer) kernel function $k : X \times X \to \R$, fulfilling
$\langle \varphi(x), \varphi(x') \rangle_{\mathcal{H}} = k(x, x')$ for
all $x, x' \in X$. Furthermore, $C > 0$ is a regularization parameter.
It controls the complexity of the SVM model.

Often, the problem is not directly solved, but dualized via Lagrangian theory.
Then the problem can be stated as 
\begin{align}
        \max_{\alpha \in \R^n} \quad & \sum_{i=1}^n \alpha_i - \frac{1}{2} \sum_{i,j=1}^n \alpha_i \alpha_j y_i y_j k(x_i, x_j) \label{SVMproblemDual} \\
        \text{s.t.} \quad & \sum_{i=1}^{n} y_i \alpha_i = 0 \text{~~and~~} 0 \leq \alpha_i \leq C \,\,\, \forall \, i \in \{1, \dots, n\}. \notag
        \enspace
\end{align}
The solution then takes the form $w = \sum_{i=1}^n \alpha_i y_i \varphi(x_i)$
and the offset~$b$ can be computed from the Karush-Kuhn-Tucker (KKT)
complementarity conditions.

SVMs classify by using the model given by
$h_{w,b}(x) = \textrm{sign}(\langle w, \phi(x)\rangle_{\mathcal{H}} + b)$

Different kernels can be used, e.g. the polynomial kernel or the 
RBF kernel $k(x,x') = e^{-\gamma || x-x' ||^2}$, where
$\gamma$ and $d$ are parameters of the kernel, that must
be tuned to the given data.

SMO.

On the practical side, LIBSVM can be regarded as the 
reference implementation of the SMO and the 'golden standard'
of SVM solvers \citep{cc01a}.

@@

Other SVM solver, like the BudgetedSVM package~\citep{djuric2014budgetedsvm}, 
which contains BSGD and LLSVM, or SVMperf~\citep{joachims2009sparse}, cannot be called directly from R.
Instead, a system command has to be issued.
To ease this calls, the \pkg{SVMBridge} package has been developed.

@@

We provided default wrappers for LIBSVM, LASVM, CVM/BVM,
SVMperf, BSGD and LLSVM.

\section{Using a Wrapper}

@@

The first step is to make the SVMBridge aware of a new wrapper.
This is simply done by calling \code{addSVMPackage()}.
There are several options.
The SVMBridge comes with an easy mechanism to search 
the corresponding binaries (specified in the wrapper) to ease
the usage. To use this, one can instead call \code{findSVMSoftware}.

@@

Most SVM solver work with the sparse data format.
This format consists of:
Each line (delimited by a CR/LF) is given by a label
and the non-zero components of the data point,
e.g. to encode the vector six dimensional vector 
 $ ( 0 , 0 , 0 , 0.2 , 0 , 1.4 ) $
belonging to class $3$, the sparse 
data format would contain the line {3 4:0.2 6:1.4}.

Although reading these files into R is possible via either
the \pkg{kernlab} or the \pkg{e1071} routines, both only provide R solutions,
and therefore suffer from suboptimal performance.
The SVMBridge package provides a simple sparse data format 
reading and writing, which are implemented in C++ and
therefore are nearly two orders of magnitudes faster than
the corresponding e1071 routines.
Notice, that currently only dense matrices are supported,
TODO: work around this with the Sparse Matrix Package.

@@

\begin{CodeChunk}
\begin{CodeInput}
R> predictions = testSVM (model, data = '')
R> head(predictions)
\end{CodeInput}
\begin{CodeOutput}
...
\end{CodeOutput}
\end{CodeChunk}

@@

\subsection{Platform Considerations}

Although the SVMBridge was meant to be cross-platform, this goal
is hard to achieve. From a users perspective, it can be used on all three
major platforms (Linux, MacOS, Windows). Additional tests should
be started to make sure the package works as intended.

@@

\section{Creating a Wrapper}

Adding your own SVM solver to the SVMBridge boils down to
writing a S3 class with several routines. In the inst folder you will find
a template that you can fill with your own code. Here we will go
through the details of adding the software package ...

\subsection{Training Call}

\subsection{Testing Call}

\subsection{Handling Models}

As only LIBSVM models are being used, one can readily use the 
inbuild functions of the SVMBridge.

@@

\bibliography{collected}
